---
title: "Reproducible Research - Project1"
author: "Lakshmi R"
date: "December 1, 2017"
output: html_document
---

## Reproducible Research - Activity Data Analysis

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Reading Data

```{r activity}
activityData <- read.csv('activity.csv')
```

# Aggergate By Day
```{r}
aggregateByDay <- aggregate(steps ~ date, activityData, sum)
```


## Histogram of the total number of steps taken each day

```{r}
hist(aggregateByDay$steps , xlab="No of Steps in a day", main="Histogram of number of steps in a day")
```

## Mean of steps taken each day

```{r}
mean(aggregateByDay$steps)
```


## Median number of steps taken each day
```{r}
median(aggregateByDay$steps)
```

## Aggregate By Interval
```{r}
averagedByInterval <- aggregate(steps ~ interval, activityData, mean)
```

# Time series plot of the average number of steps taken
```{r}
plot(averagedByInterval$interval, averagedByInterval$steps ,  xlab="Interval",
     ylab="Steps" ,main="Average number of steps taken", type="l")
```

# The 5-minute interval that, on average, contains the maximum number of steps
```{r}
averagedByInterval[which(averagedByInterval$steps==max(averagedByInterval$steps)),]$interval
```

# No of rows with Missing Value 
```{r}
nrow(activityData[(is.na(activityData$steps) |  is.na(activityData$date) |  is.na(activityData$interval)),])
```

# Filling missing value using mean of interval
```{r}
activityData_filled <- activityData
activityData_na <- is.na(activityData_filled$steps)
int_avg <- tapply(activityData_filled$steps, activityData_filled$interval, mean, na.rm=TRUE, simplify=T)
activityData_filled$steps[activityData_na] <- int_avg[as.character(activityData_filled$interval[activityData_na])]
```

## total number of steps taken each day after missing values are imputed
```{r}
aggregateByDay_filled <- aggregate(steps ~ date, activityData_filled, sum)
```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
hist(aggregateByDay_filled$steps , xlab="No of Steps in a day", main="Histogram of number of steps in a day")
```

## Mean
```{r}
mean(aggregateByDay_filled$steps)
```

## Median
```{r}
median(aggregateByDay_filled$steps)
```

## Creating new factor for dow(Day of Week)

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
activityData_filled$dow = as.factor(ifelse(is.element(weekdays(as.Date(activityData_filled$date)),weekdays), "Weekday", "Weekend"))

```
## Aggregate by DOW

```{r}
aggregateByDow <- aggregate(steps ~ interval + dow, activityData_filled, mean)

```


## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

library(lattice)
xyplot(aggregateByDow$steps ~ aggregateByDow$interval|aggregateByDow$dow, main="Average Steps across Weekends/Weekdays",xlab="Interval", ylab="Avg no of Steps",layout=c(1,2), type="l")
```








